Sequential rank agreement methods for comparison of ranked lists

### 2015-04-14 latex/paper.Rnw
Text

%
% New commands
%
\newcommand{\ie}{\emph{i.e.,}\@\xspace}
\newcommand{\eg}{\emph{e.g,}\@\xspace}

@@

Ranked lists occur in many applications of statistics. Regression
methods rank predictor variables according to magnitude of association
with outcome, prediction models rank subjects according to their risk
of an event, and genetic studies rank genes according to their
difference in expression across samples. A common research question is
where to stop, i.e., to decide the maximal significant rank.  In the
three examples this would correspond to the number of significantly
associated predictor variables, the number of patients at high risk of
an event, and the number of genes that are worth to pursue in further
experiments, respectively.

In this article we describe some new breakthrough tools for measuring
agreement across a set of lists which soon will enter the state-of-the
art. The methods should be useful whenever there are multiple rankings
of the same list. The idea is to define agreement based on the ranks
of the first k elements in each list.

@@

\subsection{All lists fully observed}
We shall start by the simplest case where all $L$ lists are fully
observed, \ie, we have the rank of all $P$ items for all of the $L$
lists. This situation occurs is common when we have the original
dataset available and when we wish to, say, compare the results from
different analysis methods.

For the fully obseved list situation we can plot the sequantial
rank agreement \eqref{def:sra} as a function of depth $d$. If there is
a ... An example is seen in Figure~\ref{XX}

@@

<<fig=TRUE, echo=FALSE>>=
library(SuperRanker)
library(glmnet)
library(minerva)
library(changepoint)
library(randomForestSRC)

@@

res <- sqrt(sra(inputmatrix))

plot(res[1:1000], lwd=3, col="red", type="l", ylim=c(0,1300), ylab="Sequential rank agreement")

## 
null <- sqrt(random_list_sra(inputmatrix, B=1, n=100))
matlines(null, col="gray", lty=1)
@

>>>>

%
% New commands
%
\newcommand{\ie}{\emph{i.e.}\@\xspace}
\newcommand{\eg}{\emph{e.g.}\@\xspace}

@@

Ranked lists occur in many applications of statistics. Regression
methods rank predictor variables according to magnitude of association
with outcome, prediction models rank subjects according to their risk
of an event, and genetic studies rank genes according to their
difference in expression across samples. A common research question is
where to stop, \ie, to decide the maximal significant rank.  In the
three examples this would correspond to the number of significantly
associated predictor variables, the number of patients at high risk of
an event, and the number of genes that are worth to pursue in further
experiments, respectively.

In this article we describe some new breakthrough tools for measuring
agreement across a set of lists which soon will enter the state-of-the
art (FIXME: denne sætning skal laves om). The methods should be useful
whenever there are multiple rankings of the same list. The idea is to
define agreement based on the ranks of the first $k$ elements in each
list.

Several other approaches have proposed to measure ... Of these, 
Kendall's $\tau$ iis the most well-known but xxx and yyy have proposed
methods that measure the ..

The manuscript is ... as follows

<<echo=FALSE>>=
#note: always pass alpha on the 0-255 scale
makeTransparent<-function(someColor, alpha=100)
{
  newColor<-col2rgb(someColor)
  apply(newColor, 2, function(curcoldata){rgb(red=curcoldata[1], green=curcoldata[2],
    blue=curcoldata[3],alpha=alpha, maxColorValue=255)})
}
@ %

@@

\subsection{All lists fully observed}
We shall start by the simplest case where all $L$ lists are fully
observed, \ie, we have the rank of all $P$ items for all of the $L$
lists. Fully observed ranked lists are common and occur, for example,
when we have a single dataset and apply different statistical analysis
methods to produce lists of predictors ranked according to their
importance, or if the same analysis method is applied to data from
different populations.

For the fully obseved list situation we can plot the sequantial
rank agreement \eqref{def:sra} as a function of depth $d$. If there is
a ... An example is seen in Figure~\ref{fig:example1}

@@

<<example1, echo=FALSE>>=
library(SuperRanker)
library(glmnet)
library(minerva)
library(changepoint)
library(randomForestSRC)

@@

res <- sqrt(sra(inputmatrix))
@

\begin{figure}[tb]
\begin{center}
<<fig=TRUE,echo=FALSE>>=
ysize <- 1000
plot(res[1:ysize], lwd=3, col="red", type="l", ylim=c(0,1300), ylab="Sequential rank agreement")

## 
# col1 <- makeTransparent()
null <- sqrt(random_list_sra(inputmatrix, B=1, n=100))
matlines(null[1:ysize,], col="gray", lty=1)

##

null2 <- sapply(1:40, function(i) {
    sra(producelists(x, sample(y)), B=1)
}
                )
matlines(null2[1:ysize,], col="lightblue", lty=1)

@ %
\end{center}
 \label{fig:example1}
 \caption{Sequential rank agreement for 4 different analysis methods
   applied to the Golub data.}
\end{figure}
